Accounting for uncertainty in simulation greenhouse gas emissions from dairy cattle in Kenya

### 2022-05-31 Manuscript_analyses.Rmd
worked on figures, mainly correcting labels and producing precise estimates (in the traditional sense) to add to Figs. 2 and 3.

```{r}
knitr::kable(read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE))

@@

```{r}
knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5])

@@

```{r plot_distributions, fig.width=10, fig.height=10}
# enteric_CH4=enteric_CH4,
enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "CH4 Emissions from Enteric Fermentation per Farm") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

# milk_yield=sum(total_milk),
milk_yield_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "milk_yield",
                   method = "boxplot_density",
                   old_names = "milk_yield",
                   new_names = "Total Milk Yield") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

# enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
enteric_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "enteric_CH4_CO2eq",
                   method = "boxplot_density",
                   old_names = "enteric_CH4_CO2eq",
                   new_names = "Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock") + 
  theme(axis.title.x=element_blank())

# mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
mm_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "mm_CH4_CO2eq",
                   method = "boxplot_density",
                   old_names = "mm_CH4_CO2eq",
                   new_names = "Carbon Dioxide Equivalent Methane Emissions from Livestock Manure") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

# mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
mm_N2O_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "mm_N2O_CO2eq",
                   method = "boxplot_density",
                   old_names = "mm_N2O_CO2eq",
                   new_names = "Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

# feed_CO2=feed_CO2,
feed_CO2_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "feed_CO2",
                   method = "boxplot_density",
                   old_names = "feed_CO2",
                   new_names = "Carbon Dioxide Emissions from Livestock Feed Production and Transport") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

# on_farm=sum(on_farm)
on_farm_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "on_farm",
                   method = "boxplot_density",
                   old_names = "on_farm",
                   new_names = "Total Carbon Dioxide Equivalent Emissions from Livestock") + 
  theme(axis.title.y=element_blank())

@@

ggsave("Fig_1_GHG_simulation_outputs.png")
```

*# THIS STILL NEEDS IMPROVEMENT - e.g. units are missing, and plot titles don't fit (use concise symbols with subscripts etc.)*

@@

plot_distributions(mcSimulation_object = farm5_emissions,
                   vars = "on_farm",
                   method = "boxplot_density",
                   old_names = "on_farm",
                   new_names = "Total Carbon Dioxide Equivalent Emissions from Livestock") + 
  theme(axis.title.y=element_blank())

ggsave("Fig_2_uncertainty_illustration.png")

@@

*# NEED TO ADD THE USUAL PRECISE VALUE HERE FOR COMPARISON (also need to add units and possibly fix up other things)*

*This*

@@

ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Emission~density~of~milk~production~(kg~CO[2]-eq~kg^-1)))

ggsave("Fig_3_emission_density_farm5.png")

@@

*# STILL NEED TO add the usual precise value estimate*

>>>>

```{r}
read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)   %>%
    kableExtra::kbl(digits = 2)  %>%
    kableExtra::kable_classic_2(full_width = F)

@@

```{r}
read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5]  %>%
    kableExtra::kbl(digits = 2)  %>%
    kableExtra::kable_classic_2(full_width = F)

We'll also make precise scenarios that contain only the exact values that were observed in the survey.

```{r}
# count the number of animals for each type

precise_estimates<-c("Ca_pasture","Ca_large_grazing","C_intact_males","C_castrates","C_females","maintenance_coefficient_Cfi_other",
                     "maintenance_coefficient_Cfi_lact_cow","maintenance_coefficient_Cfi_bulls","mature_weight_MW_female","mature_weight_MW_male",
                     "milk_fat","Cp_rate_4","Cp_rate_5","Cp_rate_6","Y_m","Bo_animaltype_1","Bo_animaltype_2","Bo_animaltype_3","Bo_animaltype_4",
                     "Bo_animaltype_5","Bo_animaltype_6","Bo_animaltype_7","Bo_animaltype_8","Bo_animaltype_9","Bo_animaltype_10","Bo_animaltype_11",
                     "Bo_animaltype_12","MCFpasture","MCFspread","MCFdrylot","MCFsolidstore","MCFcomposted","MCFliquid","MCFbiogas","MCFburn","MCFsold",
                     "mndec_pasture","mndec_spread","mndec_drylot","mndec_solidstore","mndec_composted","mndec_liquid","mnvc_pasture","mnvc_spread",
                     "mnvc_drylot","mnvc_solidstore","mnvc_composted","mnvc_liquid","frac_leach_pasture","frac_leach_spread","frac_leach_drylot",
                     "frac_leach_solidstore","frac_leach_composted","frac_leach_liquid")

precise_scenarios<-data.frame(Variable=c(variables_to_update_const,variables_to_update_posnorm,precise_estimates),param="both")

for(i in 1:nrow(animals))
  {coln<-paste0("Farm_",row.names(animals)[i])
   precise_scenarios[,coln]<-NA
     precise_scenarios[which(precise_scenarios$Variable %in% paste0("N_animal_type_",1:12)),coln]<-animals[i,]
}

for (f in Farms)
  {coln<-paste0("Farm_",f)
  # update feeding system 
  precise_scenarios[which(precise_scenarios$Variable == "feeding_system"),coln]<-
     median(Kenyaherds[which(Kenyaherds$QQNO==f),"system"])

  # update digestible energy (mean for all animals for now) (with error)
  precise_scenarios[which(precise_scenarios$Variable == "DE_perc"),coln]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"dig_energy"])

  # update crude protein (mean for all animals for now) (with error)
  precise_scenarios[which(precise_scenarios$Variable == "perc_crude_protein_diet"),coln]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"CP."])

  # update feed production CO2 (with error)
   precise_scenarios[which(precise_scenarios$Variable == "feed_prod_CO2"),coln]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])

  # update feed transport CO2 (with error)
   precise_scenarios[which(precise_scenarios$Variable == "feed_trans_CO2"),coln]<-
     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])

   # update milk yield for each cow type (animal type 4-6) (with error)
   for(typ in 4:6)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       {precise_scenarios[which(precise_scenarios$Variable==paste0("milk_yield_",typ)),coln]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "milk_yield"])
       precise_scenarios[which(precise_scenarios$Variable==paste0("CP_rate_",typ)),coln]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "Cp"]/0.1)

     }
  # update live weight and weight gain for each animal type (with error)
   for(typ in 1:12)
     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
       {precise_scenarios[which(precise_scenarios$Variable==paste0("live_weight_LW_",typ)),coln]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "live_weight_LW"])
       precise_scenarios[which(precise_scenarios$Variable==paste0("weight_gain_WG_",typ)),coln]<-
         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                         "weight_gain_WG"])
     }

   # now add the best estimates from the IPCC guidelines to the precise scenarios

   precise_scenarios[which(precise_scenarios$Variable=="Ca_pasture"),coln]<-0.17
   precise_scenarios[which(precise_scenarios$Variable=="Ca_large_grazing"),coln]<-0.36
   precise_scenarios[which(precise_scenarios$Variable=="C_intact_males"),coln]<-1.2   
   precise_scenarios[which(precise_scenarios$Variable=="C_castrates"),coln]<-1.0   
   precise_scenarios[which(precise_scenarios$Variable=="C_females"),coln]<-0.8   
   precise_scenarios[which(precise_scenarios$Variable=="maintenance_coefficient_Cfi_other"),coln]<-0.322   
   precise_scenarios[which(precise_scenarios$Variable=="maintenance_coefficient_Cfi_lact_cow"),coln]<-0.386
   precise_scenarios[which(precise_scenarios$Variable=="maintenance_coefficient_Cfi_bulls"),coln]<-0.370
   precise_scenarios[which(precise_scenarios$Variable=="mature_weight_MW_female"),coln]<-349 
   precise_scenarios[which(precise_scenarios$Variable=="mature_weight_MW_male"),coln]<-483   
   precise_scenarios[which(precise_scenarios$Variable=="milk_fat"),coln]<-4 
   precise_scenarios[which(precise_scenarios$Variable=="Cp_rate_4"),coln]<-0.1  
   precise_scenarios[which(precise_scenarios$Variable=="Cp_rate_5"),coln]<-0.1   
   precise_scenarios[which(precise_scenarios$Variable=="Cp_rate_6"),coln]<-0.1
   precise_scenarios[which(precise_scenarios$Variable=="Y_m"),coln]<-0.06

   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_1"),coln]<-0.13
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_2"),coln]<-0.13
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_3"),coln]<-0.13
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_4"),coln]<-0.13   
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_5"),coln]<-0.13   
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_6"),coln]<-0.13   
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_7"),coln]<-0.13   
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_8"),coln]<-0.13   
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_9"),coln]<-0.13   
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_10"),coln]<-0.13   
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_11"),coln]<-0.13   
   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_12"),coln]<-0.13   
   precise_scenarios[which(precise_scenarios$Variable=="MCFpasture"),coln]<-0.0047   
   precise_scenarios[which(precise_scenarios$Variable=="MCFspread"),coln]<-0.01
   precise_scenarios[which(precise_scenarios$Variable=="MCFdrylot"),coln]<-0.02     
   precise_scenarios[which(precise_scenarios$Variable=="MCFsolidstore"),coln]<-0.05
   precise_scenarios[which(precise_scenarios$Variable=="MCFcomposted"),coln]<-0.025
   precise_scenarios[which(precise_scenarios$Variable=="MCFliquid"),coln]<-0.8   
   precise_scenarios[which(precise_scenarios$Variable=="MCFbiogas"),coln]<-0.1085   
   precise_scenarios[which(precise_scenarios$Variable=="MCFburn"),coln]<-0.1   
   precise_scenarios[which(precise_scenarios$Variable=="MCFsold"),coln]<-0   
   precise_scenarios[which(precise_scenarios$Variable=="mndec_pasture"),coln]<-0.02   
   precise_scenarios[which(precise_scenarios$Variable=="mndec_spread"),coln]<-0   
   precise_scenarios[which(precise_scenarios$Variable=="mndec_drylot"),coln]<-0.02   
   precise_scenarios[which(precise_scenarios$Variable=="mndec_solidstore"),coln]<-0.01   
   precise_scenarios[which(precise_scenarios$Variable=="mndec_composted"),coln]<-0.01   
   precise_scenarios[which(precise_scenarios$Variable=="mndec_liquid"),coln]<-0.01   
   precise_scenarios[which(precise_scenarios$Variable=="mnvc_pasture"),coln]<-0.02
   precise_scenarios[which(precise_scenarios$Variable=="mnvc_spread"),coln]<-0.07
   precise_scenarios[which(precise_scenarios$Variable=="mnvc_drylot"),coln]<-0.3
   precise_scenarios[which(precise_scenarios$Variable=="mnvc_solidstore"),coln]<-0.45   
   precise_scenarios[which(precise_scenarios$Variable=="mnvc_composted"),coln]<-0.65   
   precise_scenarios[which(precise_scenarios$Variable=="mnvc_liquid"),coln]<-0.3   
   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_pasture"),coln]<-0.24   
   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_spread"),coln]<-0.24   
   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_drylot"),coln]<-0.24   
   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_solidstore"),coln]<-0.24   
   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_composted"),coln]<-0.24  
   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_liquid"),coln]<-0.24

write.csv(precise_scenarios,"data/farm_scenarios_precise.csv",row.names = FALSE)

@@

```{r plot_distributions, fig.width=10, fig.height=10}
# enteric_CH4=enteric_CH4,
enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "CH[4]~Emissions~from~Enteric~Fermentation~(kg~CH[4])") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

enteric_CH4_dist_plot <- enteric_CH4_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)

# milk_yield=sum(total_milk),
milk_yield_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "milk_yield",
                   method = "boxplot_density",
                   old_names = "milk_yield",
                   new_names = "Total~Milk~Yield~(kg)") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

milk_yield_dist_plot <- milk_yield_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)

# enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
enteric_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "enteric_CH4_CO2eq",
                   method = "boxplot_density",
                   old_names = "enteric_CH4_CO2eq",
                   new_names = "Enteric~CH[4]~Emissions~from~Livestock~(kg~CO[2]*'-eq')") + 
  theme(axis.title.x=element_blank())

enteric_CH4_CO2eq_dist_plot <- enteric_CH4_CO2eq_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)

# mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
mm_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "mm_CH4_CO2eq",
                   method = "boxplot_density",
                   old_names = "mm_CH4_CO2eq",
                   new_names = "CH[4]~Emissions~from~Livestock~Manure~(kg~CO[2]*'-eq')") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

mm_CH4_CO2eq_dist_plot <- mm_CH4_CO2eq_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)

# mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
mm_N2O_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "mm_N2O_CO2eq",
                   method = "boxplot_density",
                   old_names = "mm_N2O_CO2eq",
                   new_names = "N[2]*O~Emissions~from~Livestock~Manure~(kg~CO[2]*'-eq')") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

mm_N2O_CO2eq_dist_plot <- mm_N2O_CO2eq_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)

# feed_CO2=feed_CO2,
feed_CO2_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "feed_CO2",
                   method = "boxplot_density",
                   old_names = "feed_CO2",
                   new_names = "CO[2]~Emissions~from~Livestock~Feed~Production~and~Transport~(kg)") + 
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())
feed_CO2_dist_plot <- feed_CO2_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)

# on_farm=sum(on_farm)
on_farm_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                   vars = "on_farm",
                   method = "boxplot_density",
                   old_names = "on_farm",
                   new_names = "Total~Emissions~from~Livestock~(kg~CO[2]*'-eq')") + 
  theme(axis.title.y=element_blank())

on_farm_dist_plot <- on_farm_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)

@@

ggsave("Fig_1_GHG_simulation_outputs.png",width=9,height=9)
```

*# THIS STILL NEEDS CHECKING - are all the units correct?*

@@

precise_scenarios<-read.csv("data/farm_scenarios_precise.csv", fileEncoding = "UTF-8-BOM")

farm5_precise<-precise_scenarios[,c(1,2,7)]

farm5_emissions_precise<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
                             scenarios = farm5_precise,
                             model_function = ghg_emissions,
                             numberOfModelRuns = 10,
                             functionSyntax = "plainNames")

pd<-plot_distributions(mcSimulation_object = farm5_emissions,
                   vars = "on_farm",
                   method = "boxplot_density",
                   old_names = "on_farm",
                   new_names = "Total~Emissions~from~Livestock~(kg~CO[2]*'-eq')") + 
  theme(axis.title.y=element_blank())

pd + facet_wrap(. ~ name, labeller = label_parsed) + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=on_farm),color="blue", lwd=3)

ggsave("Fig_2_uncertainty_illustration.png",width=5,height=3)

@@

The blue line represents a classic best-bet estimate based on default values in the IPCC guidelines and some context-specific reasonable guesses.

@@

histo<-ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Emission~density~of~milk~production~(~kg~CO[2]*'-eq'~kg^-1)))

farm5_emissions_precise$y[,"CO2eq_per_milk"]<-farm5_emissions_precise$y$pc_on_farm/farm5_emissions_precise$y$pc_milk_yield

histo + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=CO2eq_per_milk),color="blue", lwd=3)

ggsave("Fig_3_emission_density_farm5.png",width=5,height=3)

@@

Also here, the blue line represents a classic best-bet estimate based on default values in the IPCC guidelines and some context-specific reasonable guesses.
